---
title: "non_eml_download_test"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries and constants

```{r libraries, warning=FALSE}
# devtools::install_github("NCEAS/metajam")
library(metajam)  

```

#### Step 2: Find the link to the dataset
Go to the web address for the dataset and find the download button for the data. 

In our case the link is: https://arcticdata.io/metacat/d1/mn/v2/object/urn%3Auuid%3A9e123f84-ce0d-4094-b898-c9e73680eafa

#### Step 3: Choose where you want the files to be saved
In our case, we'll just put it into the metajam_example folder.

```{r}

#eg desired_path_to_data <- "~/Desktop"
desired_path_to_data <- "~/metajam_test"

# create the folder if it does not exist yet
dir.create(desired_path_to_data, showWarnings = FALSE)

```

#### Step 4: Download the data by pasting the link you just copied

```{r}

my_data_url <- "https://arcticdata.io/metacat/d1/mn/v2/object/urn%3Auuid%3A9e123f84-ce0d-4094-b898-c9e73680eafa"

# this will download the data into a folder and save the path to that folder
downloaded_data <- download_d1_data(data_url = my_data_url, path = desired_path_to_data)


```

#### Step 5: Now read in the data (with all the metadata)

```{r}

my_data <- read_d1_files(downloaded_data)


```
